HO unification from object language to meta language

### 2024-02-14 main.tex
coqcode and elpicode macro for minted

\subsection{in a nutshell}
example, a rule for theorem
$$
forall f : A -> B, .. bla f .. -> prove (\forall x, f x)
bla g
$$
\begin{minted}{elpi}
type app ..
prove (forall x\ app F x) :- ... bla F ...
\end{minted}
on goal \mint{elpi}|prove (forall x\ app (app g x) x))|
would fail since \mint{elpi}|F != app g x|.

of course one wants to avoid
\begin{minted}{elpi}
prove P S S'' :- ol-unif P S (forall x\ app "F" x) S', ... assignment "F" S' F, bla F S' ...
\end{minted}

Now, ML has HO variables
\begin{minted}{elpi}
type lam ..
prove (forall x\ F x) :- ... bla F ...
\end{minted}
this time  \mint{elpi}|F x != app (app g x) x| has solution
\mint{elpi}|F = a\ app (app g x) x| but F is not a term so
bla needs to be adapted,
\begin{minted}{elpi}
prove (forall x\ F x) :- ... bla (lam F) ...
bla g.
\end{minted}
this is too simplistic since \mint{elpi}|g != lam x\ app g x|

@@

\begin{coqcode}
(* HO unif *)
Instance forall_dec A (P: A |$\to$| Prop) `{Finite A} `{|$\forall$|x, Decision (P x)} :
  Decision (|$\forall$|x, P x).
Instance and_dec P `{Decision P} Q `{Decision Q}:  Decision (P ∧ Q).

(* FO approx *)
Instance decide_rel A B (R: A |$\to$| B |$\to$| Prop) `{RelDecision R} :
  forall x y, Decision (R x y).
Instance decide_eq_nat : RelDecision (@eq nat).
Check _ : Decision (@eq nat 2 3)

@@

  Decision (|$\forall$|x, P x).

@@

\begin{minted}{elpi}
prove (app[c "decidable", all x\ P x]) Proof :- ho-link P P', prove (... P' ...).
\end{minted}

@@

questo esempio recupera l'app binaria perche

>>>>

\subsection{in a nutshell}
example, a rule for theorem
$$
forall f : A -> B, .. bla f .. -> prove (\forall x, f x)
bla g
$$
\begin{elpicode}
type app ...
prove (forall x\ app F x) :- ... bla F ...
\end{elpicode}
on goal \mint{elpi}|prove (forall x\ app (app g x) x))|
would fail since \mint{elpi}|F != app g x|.

of course one wants to avoid
\begin{elpicode}
prove P S S'' :- ol-unif P S (forall x\ app "F" x) S', ... assignment "F" S' F, bla F S' ...
\end{elpicode}

Now, ML has HO variables
\begin{elpicode}
type lam ..
prove (forall x\ F x) :- ... bla F ...
\end{elpicode}
this time  \mint{elpi}|F x != app (app g x) x| has solution
\mint{elpi}|F = a\ app (app g x) x| but F is not a term so
bla needs to be adapted,
\begin{elpicode}
prove (forall x\ F x) :- ... bla (lam F) ...
bla g.
\end{elpicode}
this is too simplistic since \mint{elpi}|g != lam x\ app g x|

@@

\begin{coqcode}
(* HO unif *)
Instance forall_dec A (P: A ~$\to$~ Prop) `{Finite A} `{~$\forall$~x, Decision (P x)} :
  Decision (~$\forall$~x, P x).
Instance and_dec P `{Decision P} Q `{Decision Q}:  Decision (P /\ Q).

(* FO approx *)
Instance decide_rel A B (R: A ~$\to$~ B ~$\to$~ Prop) `{RelDecision R} :
  forall x y, Decision (R x y).
Instance decide_eq_nat : RelDecision (@eq nat).
Check _ : Decision (@eq nat 2 3)

@@

  Decision (~$\forall$~x, P x).

@@

\begin{elpicode}
prove (app[c "decidable", all x\ P x]) Proof :- ho-link P P', prove (... P' ...).
\end{elpicode}

@@

% questo esempio recupera l'app binaria perche
